---
title: "E1- College Data Exploration"
author: "Ajit Singh"
date: "Saturday, October 18, 2014"
output: html_document
---

This exercise relates to the College data set, which can be found in the file College.csv in data folder. It contains a number of variables for 777 different universities and colleges in the US. The variables are:

- **Private** : Public/private indicator
- **Apps** : Number of applications received
- **Accept** : Number of applicants accepted
- **Enroll** : Number of new students enrolled
- **Top10perc** : New students from top 10% of high school class
- **Top25perc** : New students from top 25% of high school class
- **F.Undergrad** : Number of full-time undergraduates
- **P.Undergrad** : Number of part-time undergraduates
- **Outstate** : Out-of-state tuition
- **Room.Board** : Room and board costs
- **Books** : Estimated book costs
- **Personal** : Estimated personal spending
- **PhD** : Percent of faculty with Ph.D.'s
- **Terminal** : Percent of faculty with terminal degree
- **S.F.Ratio** : Student/faculty ratio
- **perc.alumni** : Percent of alumni who donate
- **Expend** : Instructional expenditure per student
- **Grad.Rate** : Graduation rate 

Before reading the data into R , it can be viewed in Excel or a text editor.

(a) Use the read.csv() function to read the data into R . Call the loaded data college . Make sure that you have the directory set to the correct location for the data. 
```{r}
# Provide the R code here
```

(b) Look at the data using the fix() function. You should notice that the first column is just the name of each university. We don't really want R to treat this as data. However, it may be handy to have these names for later. Try the following commands: 


```{r}
#rownames (college)=college [,1]
#fix(college)
```

You should see that there is now a row.names column with the name of each university recorded. This means that R has given each row a name corresponding to the appropriate university. R will not try to perform calculations on the row names. However, we still need to eliminate the first column in the data where the names are stored. Try

```{r}
#college=college [,-1]
#fix(college)
```

Now you should see that the first data column is Private . Note that another column labeled row.names now appears before the Private column. However, this is not a data column but rather the name that R is giving to each row.

(c) 
i. Use the **summary()** function to produce a numerical summary of the variables in the data set.
```{r}
# Provide the R code here
```

ii. Use the **pairs()** function to produce a scatterplot matrix of the first ten columns or variables of the data. Recall that you can reference the first ten columns of a matrix A using A[,1:10] .
```{r}
# Provide the R code here
```

iii. Use the **plot()** function to produce side-by-side boxplots of Outstate versus Private .
```{r}
# Provide the R code here
```
iv. Create a new qualitative variable, called **Elite** , by *binning* the **Top10perc** variable. We are going to divide universities into two groups based on whether or not the proportion of students coming from the top 10% of their high school classes exceeds 50%.

```{r}
#Elite=rep("No",nrow(college))
#Elite[college$Top10perc >50]=" Yes"
#Elite=as.factor(Elite)
#college=data.frame(college ,Elite)

```

Use the **summary()** function to see how many elite universities there are. Now use the **plot()** function to produce side-by-side boxplots of **Outstate** versus **Elite** .
```{r}
# Provide the R code here
```
v. Use the **hist()** function to produce some histograms with differing numbers of bins for a few of the quantitative variables. You may find the command **par(mfrow=c(2,2))** useful: it will divide the print window into four regions so that four plots can be made simultaneously. Modifying the arguments to this function will divide the screen in other ways.
```{r}
# Provide the R code here
```
vi. Continue exploring the data, and provide a brief summary of what you discover.
```{r}
# Provide the R code here
```